Comparing Generative Pre-trained Transformer Models to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6939 Deaths in Sierra Leone from 2019-2022

### 2025-05-24 manuscript/manuscript.tex
Rearrange data profile to methods appendix

\textbf{Methods:} 6939 records collected from VA in Sierra Leone from 2019 to 2022 were used to compare four computer models, GPT-3.5, GPT-4, InterVA-5, and InSilicoVA, to physicians for assigning CODs at the population and individual level. Open narratives were used for GPT-3.5/4 input, while structured questionnaires were used for InterVA-5/InSilicoVA input. All COD assignments were grouped into general COD categories consisting of 19, 10, and 7 categories for the adult, child, and neonatal age groups. Cause Specific Mortality Fraction (CSMF) accuracy and Partial Corrected Concordance (PCCC) were used to compare models to physicians at the population and individual level respectively. Comparisons in CSMF and PCCC to physicians among models were evaluated for all records and by COD, age group, and age ranges.

\textbf{Results:} Based on PCCC, GPT-4 had the best performance overall (0.61), followed by GPT-3.5 (0.58), InSilicoVA (0.44), and InterVA-5 (0.43). GPT-4 also had the best performance for the adult (0.65), child (0.57), and neonatal (0.62) age groups, while performance decreased as adults aged (0.72-0.59) and increased as children and neonates aged (0.5-0.7). 7 of 19, 4 of 10, and 1 of 7 adult, child, and neonatal CODs respectively had models with $\geq$0.8 PCCC. For most CODs, GPT-4 models had one of the highest performances with the exception of InSilicoVA for adult tuberculosis, child pneumonia and neonatal infections. At the population level, all models had CSMF accuracy between 0.7-0.79.

\textbf{Conclusion:} GPT and InSilicoVA models were comparable to physicians for some CODs, but require further evaluation on reliability and larger datasets.
}

@@

\keywords{Cause of Death, Physician Coding, Verbal Autopsy, GPT}

@@

VA involves two major components: survey and COD assignment \citep{worldhealthorganizationVerbalAutopsyStandards2023,chandramohanEstimatingCausesDeath2021,worldhealthorganizationVerbalAutopsyStandards2007}. In the survey component, trained lay surveyors interview those familiar with the deceased (e.g. living spouse, children, family, friends) to gather information using standardized questionnaires and open narratives. In the COD assignment component, physicians evaluate information available from the questionnaires and open narratives to assign probable CODs. This component has been criticized to be difficult to reproduce due to reliance on physician assignment \citep{gomesNationwideMortalityStudies2017,jhaProspectiveStudyOne2005,mccormickProbabilisticCauseofDeathAssignment2016b,morrisFactorsAssociatedPhysician2010,solemanVerbalAutopsyCurrent2006}. As an alternative to physician assignment, computer models, such as InterVA \citep{byassIntegratedApproachProcessing2019} and InSilicoVA \citep{mccormickProbabilisticCauseofDeathAssignment2016b}, have been studied to automatically assign CODs with performances close to physicians at the population level, but poor performances at the individual level \citep{jhaAutomatedPhysicianAssignment2019,leitaoComparisonPhysiciancertifiedVerbal2014,desaiPerformanceFourComputercoded2014,tungaVerbalAutopsyModels2021,otiVerbalAutopsyInterpretation2010}. InterVA-5 and InSilicoVA are widely used and studied standard statistical models \citep{murrayUsingVerbalAutopsy2014,otiVerbalAutopsyInterpretation2010,benaraEvaluationMethodsAssigning2023,chandramohanEstimatingCausesDeath2021,tungaVerbalAutopsyModels2021,leitaoComparisonPhysiciancertifiedVerbal2014,jhaAutomatedPhysicianAssignment2019} for COD assignment in VAs under the openVA framework \citep{liOpenVAToolkitVerbal2023}. InterVA-5 applies Bayesian probabilistic modelling \citep{bayesEssaySolvingProblem1958} using a set of standardized symptoms from reports and related conditional probabilities from medical experts to assign CODs based on the highest probability \citep{byassIntegratedApproachProcessing2019,byassStrengtheningStandardisedInterpretation2012}. InSilicoVA improves upon InterVA (e.g. comparable probabilities across individuals, measures of uncertainty, and inclusion of additional data sources) with a hierarchical Bayesian framework and Markov Chain Monte Carlo (MCMC) simulations \citep{brooksMarkovChainMonte1998,chibMarkovChainMonte2001,hanMarkovChainMonte2001} to incorporate multiple sources of uncertainty for assigning CODs based on the highest probability \citep{mccormickProbabilisticCauseofDeathAssignment2016b}. These computer models often utilize data from the structured questionnaire, but often omit the free-text open narrative, which misses latent information, such as chronology or health-seeking behaviors, that may potentially help models perform better than using the questionnaire alone \citep{jebleeAutomaticallyDeterminingCause2019,blancoExtractingCauseDeath2021,kingQualityDiagnosticValue2016}.

Recently, Large Language Models (LLM), leveraging massive datasets and deep learning approaches, have made advances in performing a variety of Natural Language Processing (NLP) tasks using free-text, such as question answering, code generation, and even medical diagnosis \citep{changSurveyEvaluationLarge2023,lundChattingChatGPTHow2023,svyatkovskiyIntelliCodeComposeCode2020,hauptAIGeneratedMedicalAdvice2023}. In 2022, a widely-available LLM called ChatGPT was released by OpenAI with capabilities of answering natural language text inquiries using training data up to September 2021. ChatGPT-3 was based on several Generative Pre-trained Transformer (GPT) models between 2018 to 2020, namely GPT-1 to GPT-3, which had notable differences in training data sizes of 5 gigabytes to 45 terabytes from web sources that resulted in 117 million to 175 billion parameter models \citep{wuBriefOverviewChatGPT2023}. In March 2023, ChatGPT-4 was released with human-level performance on various professional and academic exams and benchmarks that outperformed ChatGPT-3 \citep{openaiGPT4TechnicalReport2023}. GPT-3.5 \citep{brownLanguageModelsAre2020} and GPT-4 \citep{openaiGPT4TechnicalReport2023} are LLMs that utilize deep neural networks with transformer architectures \citep{vaswaniAttentionAllYou2017} and reinforcement learning from human feedback \citep{ouyangTrainingLanguageModels2022,christianoDeepReinforcementLearning2017,stiennonLearningSummarizeHuman2020,wirthSurveyPreferencebasedReinforcement2017} to follow instructions from prompts and provide human-level responses, with known differences in GPT-4 possessing multimodal capabilites (e.g. image/voice input/output), more recent training data, and improved responses compared to ChatGPT-3 \citep{wuBriefOverviewChatGPT2023}. Given the limited usage of free-text open narratives in computer models for determining CODs, and recent advances in LLMs that leverage natural language text prompts, we conducted a case study with Sierra Leone deaths from VA in 2019 to 2022 to compare four models, GPT-3.5, GPT-4, InterVA-5, and InSilicoVA, to physicians for determining CODs.

@@

Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \citep{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}. Thus, we used 6942 physician agreed records of the 11,920 total records. For better comparison, we standardized all codes to CGHR-10 codes (see Additional File 1) that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 to 69 years), child (28 days to 11 years), and neonatal (under 28 days) age groups. After conversion, a final total of 6939 physician agreed records (3826 adult, 2636 child, and 477 neonatal) were used for modelling and performance evaluation. See Appendix \ref{app-methods-data} and \ref{app-data} for further details on data preprocessing and characteristics of the physician agreed records.

@@

Four computer models were used to assign COD for each of the 6939 physician agreed records: GPT-3.5, GPT-4, InterVA-5, and InSilicoVA. For GPT-3.5 and GPT-4, the following user prompt was used to instruct each model to produce COD assignments as ICD-10 codes, where \verb|<age>| and \verb|<sex>| from the questionnaire, and \verb|<narrative>| from the narratives, were replaced with values from the data:

@@

\noindent For InterVA-5 and InSilicoVA, the standardized questionnaire data from EVA were converted into into OpenVA format \citep{liOpenVAToolkitVerbal2023}, before being used as input for each model to produce COD assignments as WHO VA 2016 v1.5 \citep{worldhealthorganizationVerbalAutopsyStandards2016} codes. All model outputs (ICD-10 for GPT-3.5/4 and WHO VA 2016 for InSilicoVA/InterVA-5) were converted to CGHR-10 codes to evaluate performances of models for COD assignment relative to physicians. See Appendix \ref{app-methods-model} for additional details regarding input parameters, output data, and code conversions for each model.

@@

The performance of the four models were evaluated with metrics at the population and individual level by comparing their CGHR-10 COD outputs for 6939 records. Cause Specific Mortality Fraction (CSMF) accuracy was used to evaluate models on the population level (see Appendix \ref{app-methods-csmfa}), while Partial Chance Corrected Concordance (PCCC) was used to evaluate models on the individual level (see Appendix \ref{app-methods-pccc}) \citep{murrayRobustMetricsAssessing2011}. As model performance can vary across ages and specific causes \citep{benaraEvaluationMethodsAssigning2023,murrayUsingVerbalAutopsy2014,setelValidityVerbalAutopsy2006}, the CSMF accuracy and PCCC metrics were compared for each model overall, by age group (adult, child, neonatal, by CGHR-10 COD codes, and across age ranges. For each of the adult and child age groups, metrics were calculated for five-year age ranges for records with ages at death of one-year or older and five-month age ranges for 28 days or older. For the neonatal age group, the age ranges of 0-6 days and 7-27 days were used. See Appendix \ref{app-methods-eval} for more details on performance metrics and evaluation strategy for comparing each model.

@@

Of all 6939 records, GPT-4 (0.61 PCCC) had the highest individual performance followed by GPT-3.5 (0.56 PCCC), InSilicoVA (0.44 PCCC), and InterVA-5 (0.44 PCCC) (Figure \ref{fig-pccc-csmf}). GPT-3.5 and GPT-4 had improvements ranging from 0.14-0.18 PCCC over InSilicoVA and InterVA-5, while GPT-4 slightly improved over GPT-3.5 by 0.05 PCCC. Population level performances were similar for all models (0.74-0.79 CSMF). Figure \ref{fig-pccc} shows the PCCC performance across three age groups (adult, child, and neonate). GPT-4 had the best individual performance for adult and neonatal records (0.64 and 0.58 PCCC), while GPT-3.5 had the best performance for child records (0.54 PCCC) with GPT-4 performing slightly worse (0.51 PCCC). InSilicoVA and InterVA-5 performed the worse for adult and child records ($\leq$0.5 PCCC), while GPT-3.5 performed the worse for neonatal records (0.42 PCCC). Across age ranges, all models followed a similar pattern in individual performance (Figure \ref{fig-pccc-age}). PCCC rose for records with ages from 0 to 27 days, dropped drastically for 1-5 months, rose again from 6 months to 14 years, and slowly declined for ages 15 to 69 years. The highest and lowest performances were observed for ages 10-29 years and 1-11 months respectively. Performances varied more across models for ages 0 days to 5 years, while less variation was seen between GPT-3.5 and GPT-4, as well as InSilicoVA and InterVA-5, from 5 to 69 years.

@@

The models in this study can assist physicians in assigning CODs in a variety of ways due to low costs and speed of COD assignment. Similar to differential diagnoses, GPT and InSilicoVA models offer alternative COD assignments for physicians to consider \citep{barnettComparativeAccuracyDiagnosis2019}, which can potentially help lower the number of records with ill-defined causes or reduce disagreement between physicians. At the time of this study, running GPT-3.5 cost $\sim$\$1.6 USD (\$0.5 per one million tokens), GPT-4 cost $\sim$\$115 USD (\$30 per one million tokens), and InSilicoVA was cost free on 6939 records \citep{openaiPricing2024}. These costs were lower than physicians (e.g. less than \$3 USD per house in India \citep{gomesNationwideMortalityStudies2017,jhaProspectiveStudyOne2005}), while it is possible to code over 10,000 records in under a day. When physicians are unavailable, GPT and InSilicoVA models can be a cost-efficient alternative to code large amounts of records for population estimates of CODs. However, it is recommended to apply these models only for certain CODs where models perform well, such as in Table \ref{tab-leadcod}. In addition, these models can also help divert physician resources to cases that are more difficult to code or require more attention. For example, physicians can validate cases where models performed well (e.g. maternal conditions at 0.79-0.99 PCCC), while spending more time on cases where models performed poorly (e.g. acute respiratory infections at 0.25-0.61 PCCC). Lastly, as models can assign CODs on-demand, there is potential for models to provide immediate COD estimates during the data collection process to monitor data quality.

\subsubsection{Natural Language Input and Output}\label{discuss-adv-nlf}

@@

\subsection{Output Data and Code Conversion}\label{app-methods-output}

@@

\section{Data Profile}\label{app-data}

>>>>

\textbf{Methods:} 6939 records collected from VA in Sierra Leone from 2019 to 2022 were used to compare four computer models, GPT-3.5, GPT-4, InterVA-5, and InSilicoVA, to physicians for assigning CODs at the population and individual level. Open narratives were used for GPT-3.5/4 input, while structured questionnaires were used for InterVA-5/InSilicoVA input. All COD assignments were grouped into general COD categories consisting of 19, 10, and 7 categories for adult, child, and neonatal age groups. Cause Specific Mortality Fraction (CSMF) accuracy and Partial Corrected Concordance (PCCC) were used to compare models to physicians at the population and individual level respectively. Comparisons in CSMF and PCCC to physicians among models were evaluated overall and by COD, age group, and age ranges.

\textbf{Results:} GPT-4 had the best performance overall (0.61 PCCC), followed by GPT-3.5 (0.56 PCCC), and InSilicoVA/InterVA-5 (0.44 PCCC). GPT-4 had the best performance for adult and neonatal records (0.64 and 0.58 PCCC), with GPT-3.5 for child records (0.54 PCCC). All models' performances trended upwards from 1 month to 14 years ($~$0.1-0.75 PCCC) and downwards from 15-69 years ($~$0.7-0.35) of age. GPT-4, GPT-3.5, ad InSilicoVA had the highest performances for 17, 9, and 4 of all 30 CODs respectively. At the population level, all models had CSMF accuracies between 0.7-0.79.

\textbf{Conclusion:} All models performed well at the population level, while GPT-3.5/4 and InSilicoVA performed well for some CODs. GPT models are unable to replace physician coding yet, but have made large improvements over InterVA-5, setting the foundation for assisting physicians with alternative diagnoses to reduce ill-defined codes and disagreement when coding VAs.
}

@@

\keywords{Cause of Death, Physician Coding, Verbal Autopsy, GPT, AI, LLM}

@@

VA involves two major components: survey and COD assignment \citep{worldhealthorganizationVerbalAutopsyStandards2023,chandramohanEstimatingCausesDeath2021,worldhealthorganizationVerbalAutopsyStandards2007}. In the survey component, trained lay surveyors interview those familiar with the deceased (e.g. living spouse, children, family, friends) to gather information using standardized questionnaires and open narratives. In the COD assignment component, physicians evaluate information available from the questionnaires and open narratives to assign probable CODs. This component has been criticized to be difficult to reproduce due to reliance on physician assignment \citep{gomesNationwideMortalityStudies2017,jhaProspectiveStudyOne2005,mccormickProbabilisticCauseofDeathAssignment2016b,morrisFactorsAssociatedPhysician2010,solemanVerbalAutopsyCurrent2006}. As an alternative to physician assignment, computer models, such as InterVA \citep{byassIntegratedApproachProcessing2019} and InSilicoVA \citep{mccormickProbabilisticCauseofDeathAssignment2016b}, have been studied to automatically assign CODs with performances close to physicians at the population level, but poor performances at the individual level \citep{jhaAutomatedPhysicianAssignment2019,leitaoComparisonPhysiciancertifiedVerbal2014,desaiPerformanceFourComputercoded2014,tungaVerbalAutopsyModels2021,otiVerbalAutopsyInterpretation2010}. These computer models often utilize data from the structured questionnaire, but often omit the free-text open narrative, which misses latent information, such as chronology or health-seeking behaviors, that may potentially help models perform better than using the questionnaire alone \citep{jebleeAutomaticallyDeterminingCause2019,blancoExtractingCauseDeath2021,kingQualityDiagnosticValue2016}.

Recently, Large Language Models (LLM), leveraging massive datasets and deep learning approaches, have made advances in performing a variety of Natural Language Processing (NLP) tasks using free-text, such as question answering, code generation, and even medical diagnosis \citep{changSurveyEvaluationLarge2023,lundChattingChatGPTHow2023,svyatkovskiyIntelliCodeComposeCode2020,hauptAIGeneratedMedicalAdvice2023}. In 2022, a widely-available LLM called ChatGPT was released by OpenAI with capabilities of answering natural language text inquiries using training data up to September 2021. ChatGPT-3 was based on several Generative Pre-trained Transformer (GPT) models between 2018 to 2020, namely GPT-1 to GPT-3, which had notable differences in training data sizes of 5 gigabytes to 45 terabytes from web sources that resulted in 117 million to 175 billion parameter models \citep{wuBriefOverviewChatGPT2023}. In March 2023, ChatGPT-4 was released with human-level performance on various professional and academic exams and benchmarks that outperformed ChatGPT-3 \citep{openaiGPT4TechnicalReport2023}. Given the limited usage of free-text open narratives in computer models for determining CODs, and recent advances in LLMs that leverage natural language text prompts, we conducted a case study with Sierra Leone deaths from VA in 2019 to 2022 to compare four models, GPT-3.5, GPT-4, InterVA-5, and InSilicoVA, to physicians for determining CODs.

@@

Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \citep{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}. Thus, we used 6942 physician agreed records of the 11,920 total records. For better comparison, we standardized all codes to CGHR-10 codes (see Additional File 1) that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 to 69 years), child (28 days to 11 years), and neonatal (under 28 days) age groups. After conversion, a final total of 6939 physician agreed records (3826 adult, 2636 child, and 477 neonatal) were used for modelling and performance evaluation. See Appendix \ref{app-methods-data} for further details on data preprocessing and characteristics of the physician agreed records.

@@

Four computer models were used to assign COD for each of the 6939 physician agreed records: GPT-3.5, GPT-4, InterVA-5, and InSilicoVA. InterVA-5 and InSilicoVA are widely used and studied standard statistical models \citep{murrayUsingVerbalAutopsy2014,otiVerbalAutopsyInterpretation2010,benaraEvaluationMethodsAssigning2023,chandramohanEstimatingCausesDeath2021,tungaVerbalAutopsyModels2021,leitaoComparisonPhysiciancertifiedVerbal2014,jhaAutomatedPhysicianAssignment2019} for COD assignment in VAs under the openVA framework \citep{liOpenVAToolkitVerbal2023}. InterVA-5 applies Bayesian probabilistic modelling \citep{bayesEssaySolvingProblem1958} using a set of standardized symptoms from reports and related conditional probabilities from medical experts to assign CODs based on the highest probability \citep{byassIntegratedApproachProcessing2019,byassStrengtheningStandardisedInterpretation2012}. InSilicoVA improves upon InterVA (e.g. comparable probabilities across individuals, measures of uncertainty, and inclusion of additional data sources) with a hierarchical Bayesian framework and Markov Chain Monte Carlo (MCMC) simulations \citep{brooksMarkovChainMonte1998,chibMarkovChainMonte2001,hanMarkovChainMonte2001} to incorporate multiple sources of uncertainty for assigning CODs based on the highest probability \citep{mccormickProbabilisticCauseofDeathAssignment2016b}. GPT-3.5 \citep{brownLanguageModelsAre2020} and GPT-4 \citep{openaiGPT4TechnicalReport2023} are LLMs that utilize deep neural networks with transformer architectures \citep{vaswaniAttentionAllYou2017} and reinforcement learning from human feedback \citep{ouyangTrainingLanguageModels2022,christianoDeepReinforcementLearning2017,stiennonLearningSummarizeHuman2020,wirthSurveyPreferencebasedReinforcement2017} to follow instructions from prompts and provide human-level responses, with known differences in GPT-4 possessing multimodal capabilites (e.g. image/voice input/output), more recent training data, and improved responses compared to ChatGPT-3 \citep{wuBriefOverviewChatGPT2023}.

For GPT-3.5 and GPT-4, the following user prompt was used to instruct each model to produce COD assignments as ICD-10 codes, where \verb|<age>| and \verb|<sex>| from the questionnaire, and \verb|<narrative>| from the narratives, were replaced with values from the data:

@@

\noindent For InterVA-5 and InSilicoVA, the standardized questionnaire data from EVA were converted into OpenVA format \citep{liOpenVAToolkitVerbal2023}, before being used as input for each model to produce COD assignments as WHO VA 2016 codes \citep{worldhealthorganizationVerbalAutopsyStandards2016}. All model outputs were converted to CGHR-10 codes to evaluate performances of models for COD assignment relative to physicians. See Appendix \ref{app-methods-model} for additional details regarding input parameters, output data, and code conversions for each model.

@@

The performance of the four models were evaluated with metrics at the population and individual level by comparing their CGHR-10 COD outputs for 6939 records. Cause Specific Mortality Fraction (CSMF) accuracy was used to evaluate models on the population level (see Appendix \ref{app-methods-csmfa}), while Partial Chance Corrected Concordance (PCCC) was used to evaluate models on the individual level (see Appendix \ref{app-methods-pccc}) \citep{murrayRobustMetricsAssessing2011}. As model performance can vary across ages and specific causes \citep{benaraEvaluationMethodsAssigning2023,murrayUsingVerbalAutopsy2014,setelValidityVerbalAutopsy2006}, the CSMF accuracy and PCCC metrics were compared for each model overall, by age group (adult, child, neonatal), by CGHR-10 COD codes, and across age ranges. For each of the adult and child age groups, metrics were calculated for five-year age ranges for records with ages at death of one-year or older and five-month age ranges for 28 days or older. For the neonatal age group, the age ranges of 0-6 days and 7-27 days were used. See Appendix \ref{app-methods-eval} for more details on performance metrics and evaluation strategy for comparing each model.

@@

Of all 6939 records, GPT-4 (0.61 PCCC) had the highest individual performance followed by GPT-3.5 (0.56 PCCC), InSilicoVA (0.44 PCCC), and InterVA-5 (0.44 PCCC) (Figure \ref{fig-pccc-csmf}). GPT-3.5 and GPT-4 had improvements ranging from 0.14-0.18 PCCC over InSilicoVA and InterVA-5, while GPT-4 slightly improved over GPT-3.5 by 0.05 PCCC. Population level performances were similar for all models (0.74-0.79 CSMF). Figure \ref{fig-pccc} shows the PCCC performance across three age groups (adult, child, and neonate). GPT-4 had the best individual performance for adult and neonatal records (0.64 and 0.58 PCCC), while GPT-3.5 had the best performance for child records (0.54 PCCC) with GPT-4 performing slightly worse (0.51 PCCC). InSilicoVA and InterVA-5 performed the worse for adult and child records ($\leq$0.5 PCCC), while GPT-3.5 performed the worse for neonatal records (0.42 PCCC). Across age ranges, all models followed a similar pattern in individual performance (Figure \ref{fig-pccc-age}). PCCC trended upwards for 1 month to 14 years ($~$0.1-0.75), and downwards for ages 15 to 69 years ($~$0.7-0.35). The highest and lowest performances were observed for ages 10-29 years ($~$) and 1-11 months respectively. Performances varied more across models for ages 0 days to 5 years, while less variation was seen between GPT-3.5 and GPT-4, as well as InSilicoVA and InterVA-5, from 5 to 69 years.

@@

The models in this study can assist physicians in assigning CODs in a variety of ways due to low costs and speed of COD assignment. Similar to differential diagnoses, GPT and InSilicoVA models offer alternative COD assignments for physicians to consider \citep{barnettComparativeAccuracyDiagnosis2019}, which can potentially help lower the number of records with ill-defined causes or reduce disagreement between physicians. At the time of this study, running GPT-3.5 cost $\sim$\$1.6 USD (\$0.5 per one million tokens), GPT-4 cost $\sim$\$115 USD (\$30 per one million tokens), and InSilicoVA was cost free on 6939 records \citep{openaiPricing2024}. These costs were lower than physicians (e.g. less than \$3 USD per house in India \citep{gomesNationwideMortalityStudies2017,jhaProspectiveStudyOne2005}), while it is possible to code over 10,000 records in under a day. When physicians are unavailable, GPT and InSilicoVA models can be a cost-efficient alternative to code large amounts of records for population estimates of CODs. However, it is recommended to apply these models only for certain CODs where models perform well, such as in Table \ref{tab-leadcod}. In addition, these models can also help divert physician resources to cases that are more difficult to code or require more attention. For example, physicians can validate cases where models performed well (e.g. maternal conditions at 0.79-0.99 PCCC), while spending more time on cases where models performed poorly (e.g. acute respiratory infections at 0.25-0.61 PCCC). Lastly, as models can assign CODs on-demand, there is potential for models to provide immediate COD estimates during the data collection process to monitor data quality (e.g. comparing estimated to expected COD distributions for known areas as quality checks).

\subsubsection{Natural Language Input and Output}\label{discuss-adv-nl}

@@

\subsubsection{Output Data and Code Conversion}\label{app-methods-output}

@@
